Cross Validation Approaches for Penalized Cox Regression

### 2019-05-08 main.tex
update data analysis section

\subsection {Cross-Validated C Index}
\bnote{unified the naming of C index in this manuscript. Use Harrell's C Index or C Index instead of other alternatives. This is the default concordance statistic that the survival package would return.}

@@

\bnote{(END)}

@@

In this section, we apply the cross validation methods described in Section~\ref{Sec:methods} to data from a study of gene expression and survival in patients with lung cancer. \citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
They gathered expression profiles of \textit{22,283} genes for \textit{442} \note{is there some reason 442 is in italics? other numbers aren't} non-small-cell lung cancer (NSCLC) patients along with a number of clinical variables such as stage, age and gender. \note{Did you only adjust for stage, age, and gender, or did you adjust for smoking, etc., as well?  If not, maybe take out ``such as'', which implies that there are other variables you're not mentioning.} About 50\% of the patients died during the course of the study (236 events), with the rest censored. The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.

For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}.  We used cross-validation to select the regularization parameter $\lam$ as well as the elastic net tuning parameter $\alpha$, where $\alpha=1$ corresponds to the lasso and $\alpha=0$ corresponds to ridge regression.  \note{More on elastic net, justify why we focus on $\alpha=0.3$}.

\note{This paragraph will probably need to change based on new results.}  How the CV methods perform on real data reflects what is shown in simulation studies. The standard CVE is the most liberal approach. Deviance Residual and Verweij and Van Houwelingen's CVE approaches were both really conversative, while the standard approach was the second liberal to standard CVE. In Figure 5, the Cross Validated Error is rescaled and plotted for all four methods. A $\lambda$ will be selected when CVE curve reaches its lower point. The blue line, which represents the linear predictor approach, has more curvature near its lowest point. It is easier to pick out the minimum point for this blue curve. Hence this approach is better at picking out signals than the other three approaches. For Verweij and Van Houwelingen's cross-validated partial likelihood, there's almost no signal there.

\begin{figure}[h]
  \centering
  \includegraphics[height= 8cm ]{./figures/shedden.jpg}
  \caption{Comparing Cross-Validated Error for Different Methods on Real Data Sets \note{Change to $\alpha=0.3$?}}
\end{figure}

\begin{table}[h]
\centering
\caption{Application to Real Data Sets\note{Change to $\alpha=0.3$?}}
\label{my-label}
\begin{tabular}{lccccc}
  \toprule
  & $\lambda$ & log($\lambda$) & CV C-Index & Brier Score \\ 
						\cline{2-5} \\ 
						V $\&$ VH& 0.123 & -2.097 & 0.640 & 0.246\\
						standard & 0.102 & -2.278 & 0.643& 0.239\\
						Linear Pred &0.085 & -2.460 & 0.651&  0.227\\
						Dev Resid & 0.127& -2.066 & 0.639& 0.246\\ 
						\bottomrule
\end{tabular}
\end{table}

@@

\section{Discussion}

>>>>

\subsection {Cross-Validated C Index}
\label{Sec:CIndex}
\bnote{unified the naming of C index in this manuscript. Use Harrell's C Index or C Index instead of other alternatives. This is the default concordance statistic that the survival package would return.}

@@

@@

In this section, we demonstrate the usage and performance of the aforementioned cross validation methods when fitting penalized regression models on data from a study of gene expression and survival in patients with lung cancer. \citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
They gathered expression profiles of 22,283 genes for 442 non-small-cell lung cancer (NSCLC) patients along with some additional clinical and demographic data. About $50\%$ of the patients died during the course of the study (236 events), with the rest censored. The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.

\begin{table}[h]
\centering
\begin{tabular}{ccccccccccc}
\toprule
 & \multicolumn{2}{c}{$\alpha = 0.3$ } & \multicolumn{2}{c}{$\alpha = 0.5$} & \multicolumn{2}{c}{$\alpha = 0.7$} & \multicolumn{2}{c}{$\alpha = 0.9$} & \multicolumn{2}{c}{$\alpha = 1$} \\
 &      $\lambda$      &   CVE         &   $\lambda$        &     CVE       &      $\lambda$      &     CVE       &    $\lambda$        &    CVE        &     $\lambda$      &    CVE       \\

\hline\\

 V$\&$VH  & 0.363 &  296.32  &  0.228   &  296.35  & 0.163 &  296.34  &  0.133  &   296.31  &  0.119  &  296.30  \\
 Standard & 0.301 & 145.17  & 0.218  &  145.38 & 0.163  & 145.37 &  0.127 &  145.35 & 0.114   & 145.34 \\
 Linear Pred & 0.288 & 10.661  & 0.181  & 10.673 & 0.163 & 10.673    & 0.127  & 10.672  &    0.114 &  10.672  \\
 Dev Resid & 0.316 & 571.40 & 0.228  &  571.96 & 0.163 & 571.95  & 0.133  &   571.79 &  0.119  &  571.70  \\
\bottomrule
\end{tabular}
\caption{\bnote{Not sure whether to include this table or not}}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{lcccccc}
  \toprule
  & $\lambda$ & log($\lambda$) & $\#$ of Genes Selected & CV C-Index & Brier Score \\ 
\cline{2-6} \\ 
V $\&$ VH& 0.363 & -1.013 &  & 0.639 & 0.243\\
standard & 0.301 & -1.200 &  & 0.647& 0.232\\
Linear Pred &0.288 & -1.246 &  & 0.650&  0.229\\
Dev Resid & 0.316 & -1.153 &  & 0.645& 0.235\\ 
\bottomrule
\end{tabular}
\caption{\label{tb:shedden}Results of elasticnet model with $\alpha = 0.3$ on the Lung Cancer Data Set. The linear Predictor approach picked the smallest $\lambda$. The corresponding model yields the highest CV C-Index and smallest Brier Score evaluated at median survvial time \bnote{I got different number of genes for LP approach...}}
\end{table}

For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}. Besides the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression. For each $\alpha = 0.3, 0.5, 0.7, 0.9 \text{ and } 1$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error. The standard approach, linear predictor approach and deviance approach all reach the lowest cross-validation error at $\alpha = 0.3$.

We report the predictive performance for all four models at $\alpha = 0.3$ and their corresponding $\lambda$s in Table~\ref{tb:shedden}. For each model, we computed the cross-validated C-Index and Brier Scores at predicting the median survival time. How the CV methods perform on real data reflects what is shown in simulation studies. The standard approach and the linear predictor approach, selecting smaller $\lambda$ values, were the more liberal, while Verweij and Van Houwelingen's CVE approaches were both more conversative. With the highest C Index and lowest Brier Score, the linear predictor approach has best predictive performance among all four methods. The same patterns are repeated when $\alpha$ is at different values.

We compared the CVE along the solution path for different methods. In Figure~\ref{fig:CVE}, the Cross Validated Error is rescaled and plotted for all four methods. A $\lambda$ is selected when CVE curve reaches its lower point. The blue line and green line, which represent the linear predictor approach and standard approah, have more curvature near theirs lowest point, making it is easier to pick out the minimum point.

\begin{figure}[h]
  \centering
  \includegraphics[height= 8cm ]{./manuscript_figure/shedden.jpg}
  \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$. The horizonal axis is $\log(\lambda)$. The vertical axis is the cross-validated error rescaled to a range from 0 to 1. After the rescaling, the Cross Validated Error of four different methods can be plotted together for comparison. }
\end{figure}

@@

\bnote{(END)}
\section{Discussion}
